---
title: "PA1-template"
author: "Anna"
date: "7/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning=FALSE)
```

## Loading and preprocessing the data

```{r}
library(ggplot2)
activity <- read.csv("~/OneDrive - Johns Hopkins University/R/activity.csv")
activity$date <- as.POSIXct(activity$date, format="%Y-%m-%d")
weekday <- weekdays(activity$date)
activity <- cbind(activity,weekday)
summary(activity)
```

## 1. What is mean total number of steps taken per day?

```{r, echo=TRUE}
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
hist(activity_total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "darkblue", ylim = c(0,20), breaks = seq(0,25000, by=2500))
```

Here is the mean of the total number of steps taken per day:

```{r}
mean(activity_total_steps$steps)
```
Here is the median of the total number of steps taken per day:
```{r}
median(activity_total_steps$steps)
```

## 2. What is the average daily activity pattern?

Make a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
avg_step <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
plot(avg_step$interval, avg_step$steps, type = "l", lwd = 2, col = "darkblue",
     main = "Time Series: Average Number of Steps Taken", axes = FALSE,
     xlab = "5-minute interval", ylab = "Average number of steps")
axis(1)
axis(2, las = 1)
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
avg_step$interval[which.max(avg_step$steps)]
```

## 3. Imputing missing values

```{r}
sum(is.na(activity$steps))
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
imp <- activity # new dataset called imp
for (i in avg_step$interval) {
    imp[imp$interval == i & is.na(imp$steps), ]$steps <- 
        avg_step$steps[avg_step$interval == i]
}
head(imp) # no NAs
sum(is.na(imp)) # should be 0
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
total_step_imp <- aggregate(steps ~ date, data = imp, sum, na.rm = TRUE)
hist(total_step_imp$steps, breaks = 20, 
     main = "Total Number of Steps Taken Each Day (Imputed)",
     col = "darkblue", border = "white", xlab = "Step", axes = FALSE)
axis(1)
axis(2, las = 1)
```

Here is the mean of the total number of steps taken per day:

```{r}
mean(total_step_imp$steps)
```
Here is the median of the total number of steps taken per day:
```{r}
median(total_step_imp$steps)
```

## 4. Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
imp$day <- weekdays(imp$date)
imp$week <- ""
imp[imp$day == "Saturday" | imp$day == "Sunday", ]$week <- "weekend"
imp[!(imp$day == "Saturday" | imp$day == "Sunday"), ]$week <- "weekday"
imp$week <- factor(imp$week)
```

Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
avg_step_imp <- aggregate(steps ~ interval + week, data = imp, mean)
library(lattice)
xyplot(steps ~ interval | week, data = avg_step_imp, type = "l", lwd = 2,
       layout = c(1, 2), 
       xlab = "5-minute interval", 
       ylab = "Average number of steps",
       main = "Average Number of Steps Taken (across all weekday days or weekend days)")
```




